A data-driven, pixel-level model to improve the precision of \Kepler\ photometry

### 2013-10-15 documents/plm.tex
figure one; tweaks, notes to self

\newcommand{\notenglish}[1]{\textit{#1}}
\newcommand{\sic}{\notenglish{sic}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\Kepler}{\project{Kepler}}
\newcommand{\name}{PLM}

@@

In general, there are many reasons for apparent photometric variability in a \Kepler\ source.
There is intrinsic stellar variability,
  which is of interest to some and a nuisance to others.
There is also variability of overlapping fainter stars;
  that is, confusion noise combined with variability of the confusing sources.
There are small changes in spacecraft pointing,
  which leads to slightly different illumination of the focal-plane pixels,
  and thus different sensitivity to errors (problems) in the device flat-field or sensitivity map.
There are also \emph{intra-pixel} sensitivity variations that can contribute (CITE WHITEPAPER, SPITZER).
There are small changes in spacecraft temperature,
  which lead to point-spread function (PSF) and differential (across the focal plane) pointing changes.
These also lead to changes in pixel and intra-pixel illumination.
Stellar proper motion, geometric parallaxes, and differential stellar aberration as the spacecraft orbits all do more of the same.
There is electronic cross-talk between detectors and charge-transfer inefficiency;
  these can effectively transfer variability from one source to another.
There are additional electronics effects like ``rolling bands'' that put additional features into lightcurves.
There are also probably changes to the detector sensitivity with temperature and time,
  possibly illimunation-history effects,
  and possibly sources of variability not yet considered.
The remarkable thing about \Kepler\ is that it is trying to measure stars at a level of precision
  much higher than ever previously attempted;
  new effects really \emph{must} appear at some point.
In \figurename~HOGG, we show the pixel-level variability in the \Kepler\ data
  near one bright star that shows minimal intrinsic variability;
  this figure highlights the spacecraft-induced effects.

@@

\section{Generalities: Data-driven models}

@@

In the context of this project, the quantities of interest are the light-curves
  (photometry as a function of time or apparent flux as a function of time)
  for every star.
What we get, for every star, from the data-driven model, is a train-and-test prediction
  for the brightness of every pixel in the patch of the detector relevant to the star,
  along with the residual, or the difference between the measured read-out pixel value and the prediction (data minus model).
What will be called the ``\name\ photometry'' in what follows is constructed from what will be called the ``\name\ images''.
The \name\ photometry is constructed from the \name\ images just as the \Kepler\ official SAP photometry is constructed from the XXX images;
  that is, by summing pre-defined sets of pixels (with unit weights; see above and below for criticism of this).
The \name\ images are constructed from the data-driven model predictions by taking,
  for each pixel in each contiguous data segment (see below),
  the \emph{mean} predicted pixel value (mean over time),
  and adding to it the residuals (data minus model) away from the data-driven model prediction.
These \name\ images are ``calibrated'' images,
  in that the data-driven model has been used to fit out all the variability it can,
  subject to the restrictions imposed by its finite freedom and the train-and-test specification.
The signal---in which exoplanet transits will be found---is fundamentally the observed residual away from this prediction,
  but it has to be added back into the mean prediction to have the right ``baseline'' or zero level.
In short, \name\ will produce prediction images, residual images, \name\ images (mean prediction plus residual),
  and aperture-based \name\ photometry analogous to the SAP photometry.

@@

\section{Specifics: \name\ specification and hyper-parameters}

@@

Not the least of these is linearity.

Interesting whether the output of the model can be interrogated to provide value for those building physical models.

@@

WHITEPAPER AUTHORS:  FEEL FREE TO CONTRIBUTE TO THIS PAPER AND MOVE FROM THE ACK LIST TO THE AUTHOR LIST.

>>>>

\newcommand{\notenglish}[1]{\textit{#1}}
\newcommand{\sic}{\notenglish{sic}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\Kepler}{\project{Kepler}}
\newcommand{\name}{PLM}
\newcommand{\todo}[1]{\textbf{#1}}

@@

In general, there are many reasons for apparent photometric variability in a \Kepler\ source.
There is intrinsic stellar variability,
  which is of interest to some and a nuisance to others.
There is also variability of overlapping fainter stars;
  that is, confusion noise combined with variability of the confusing sources.
There are small changes in spacecraft pointing,
  which leads to slightly different illumination of the focal-plane pixels,
  and thus different sensitivity to errors (problems) in the device flat-field or sensitivity map.
There are also \emph{intra-pixel} sensitivity variations that can contribute (CITE WHITEPAPER, SPITZER).
There are small changes in spacecraft temperature,
  which lead to point-spread function (PSF) and differential (across the focal plane) pointing changes.
These also lead to changes in pixel and intra-pixel illumination.
Stellar proper motion, geometric parallaxes, and differential stellar aberration as the spacecraft orbits all do more of the same.
There is electronic cross-talk between detectors and charge-transfer inefficiency;
  these can effectively transfer variability from one source to another.
There are additional electronics effects like ``rolling bands'' that put additional features into lightcurves.
There are also probably changes to the detector sensitivity with temperature and time,
  possibly illimunation-history effects,
  and possibly sources of variability not yet considered.
The remarkable thing about \Kepler\ is that it is trying to measure stars at a level of precision
  much higher than ever previously attempted;
  new effects really \emph{must} appear at some point.
In \figurename~\ref{fig:pixelpatch}, we show the pixel-level variability in the \Kepler\ data
  near one bright star that shows minimal intrinsic variability;
  this figure highlights the spacecraft-induced effects.

@@

\section{Data-driven models}

@@

In the context of this project, the quantities of interest are the light-curves
  (photometry as a function of time or apparent flux as a function of time)
  for every star.
What we get, for every star, from the data-driven model, is a train-and-test prediction
  for the brightness of every pixel in the patch of the detector relevant to the star,
  along with the residual, or the difference between the measured read-out pixel value and the prediction (data minus model).
What will be called the ``\name\ photometry'' in what follows is constructed from what will be called the ``\name\ images''.
The \name\ photometry is constructed from the \name\ images
  just as the \Kepler\ official SAP photometry is constructed from the \todo{XXX images};
  that is, by summing pre-defined sets of pixels (with unit weights; see above and below for criticism of this).
The \name\ images are constructed from the data-driven model predictions by taking,
  for each pixel in each contiguous data segment (see below),
  the \emph{mean} predicted pixel value (mean over time),
  and adding to it the residuals (data minus model) away from the data-driven model prediction.
These \name\ images are ``calibrated'' images,
  in that the data-driven model has been used to fit out all the variability it can,
  subject to the restrictions imposed by its finite freedom and the train-and-test specification.
The signal---in which exoplanet transits will be found---is fundamentally the observed residual away from this prediction,
  but it has to be added back into the mean prediction to have the right ``baseline'' or zero level.
In short, \name\ will produce prediction images, residual images, \name\ images (mean prediction plus residual),
  and aperture-based \name\ photometry analogous to the SAP photometry.

@@

\section{\name\ specification and hyper-parameters}

@@

Not the least of these is linearity; deep learning and so on.

The weakest point is construction of photometry from the model plus residuals;
  interesting whether the output of the model can be interrogated
  to provide value for those building physical models.

@@

\todo{White-paper Authors:  Feel free to contribute to this paper and move from the ack list to the author list.}

\clearpage
\begin{figure}
~ % DFM: replace this ~ with a figure
\caption{A patch of \Kepler\ imaging from \todo{XXX quarter}, centered on \todo{YYY target}.
  In the top panel, each box represents a pixel.
  Within each box is shown the light-curve of that pixel over the quarter.
  In the bottom panel, the SAP photometry for this target is shown;
  the SAP photometry is based on a straight sum of the pixels marked with darker borders.
  Note that every pixel varies substantially but the total SAP photometry is better than any individual pixel.
  Note also that the pixel variations are strongly inter-related.\label{fig:pixelpatch}}
\end{figure}
